The Astropy Project

### 2017-11-25 main.tex

The development of the \astropypkg core package began as a largely
community-driven effort to standardize core functionality for astronomical
software in \python.
In this way, its genesis differs from but builds upon many substantial and
% * <bsipocz@gmail.com> 2017-11-13T13:13:08.884Z:
%
% how do astroy builds upon these? or it's more like builds upon the developer experience?
%
%
% ^.
former astronomical software development efforts that were commissioned or
initiated through large institutional support, for example IRAF \citep[developed
at NOAO;][]{IRAF}, MIDAS \citep[developed at ESO;][]{MIDAS}, or Starlink
\citep[originally developed by a consortium of UK institutions and now
maintained by the East Asian Observatory;][]{starlink1982,starlink2013}.
More recently, community-driven efforts have seen significant success in the astronomical sciences \citep{yt}.

\python\footnote{\url{https://www.python.org/}} is an increasingly popular, general-purpose
programming language that is available under a permissive open source software license free of
charge for all major operating systems. The programming language has become especially popular
in the quantitative sciences, where researchers must simultaneously produce research, perform 
data analysis and develop software. A large part of this success owes itself to the vibrant 
community of developers and a continuously-growing ecosystem of tools, web services and stable
well-developed packages that enable easier collaboration on software development, easier 
writing and sharing of software documentation and continuous testing and validation of 
software. While the programming language provides support for array representation \& 
arithmetic \citep[\package{numpy};][]{numpy}, a wide variety of functions for scientific 
computing \citep[\package{scipy};][]{numpy}, publication-quality plotting 
\citep[\package{matplotlib};][]{matplotlib}, tens of thousands of other high quality and easy 
to use packages are available which can help with tasks that are not astronomy specific but 
might be performed in the course of astronomical research, for example, interfacing with 
databases or statistical inferences. More recently, the development of package managers such as
Anaconda\footnote{\url{https://anaconda.org/}} have streamlined the installation process for
most packages significantly reducing the barrier to entry for using many such libraries.

@@

\begin{figure}
\includegraphics[width=\textwidth]{bayesian_blocks_hist.pdf}
\caption{%
    A histogram with irregular bins, fit using the Bayesian blocks
    algorithm. The left panel shows a standard histogram using
    \package{matplotlib}'s default of 10 bins. The right panel shows
    the same histogram with bin widths determined by the Bayesian
    blocks model. The model is such that relevant features of the data
    distribution are much more apparent.
    \label{fig:bayes-blocks-hist}
}
\end{figure}

>>>>

The development of the \astropypkg core package began as a largely
community-driven effort to standardize core functionality for astronomical
software in \python.
In this way, its genesis differs from but builds upon many substantial and
% * <bsipocz@gmail.com> 2017-11-13T13:13:08.884Z:
%
% how do astroy builds upon these? or it's more like builds upon the developer experience?
%
%
% ^.
former astronomical software development efforts that were commissioned or
initiated through large institutional support, for example IRAF \citep[developed
at NOAO;][]{IRAF}, MIDAS \citep[developed at ESO;][]{MIDAS}, or Starlink
\citep[originally developed by a consortium of UK institutions and now
maintained by the East Asian Observatory;][]{starlink1982,starlink2013}.
More recently, community-driven efforts have seen significant success in the astronomical sciences \citep{yt}.
\python\footnote{\url{https://www.python.org/}} is an increasingly popular,
general-purpose programming language that is available under a permissive open
source software license free of charge for all major operating systems.
Stable and well-developed packages provide support for array representation and arithmetic
\citep[\package{numpy};][]{numpy}, a wide variety of functions for scientific
computing \citep[\package{scipy};][]{numpy}, and publication-quality plotting
\citep[\package{matplotlib};][]{matplotlib}. Tens of thousands of other packages
are available which can help with tasks that are not astronomy specific but
might be performed in the course of astronomical research, for example interfacing with databases or statistical inference.

\python has become especially popular in the quantitative sciences,
where researchers must simultaneously produce research, perform data analysis,
and develop software.
A large part of this success owes itself to a vibrant community of developers
and a continuously-growing ecosystem of tools and web services that enable
easier collaboration on software development, easier writing and sharing of
software documentation, and continuous testing and validation of software.
More recently, package managers such as
Anaconda\footnote{\url{https://anaconda.org/}} have streamlined the installation
process for most packages, therefore significantly reducing the barrier to entry
for using many such libraries.

@@

\begin{figure}
\includegraphics[width=\textwidth]{bayesian_blocks_hist.pdf}
\caption{%
    Three approaches to a 1D histogram:
    {\it left:} a standard histogram using \package{matplotlib}'s default of 10 bins.
    {\it center:} a histogram with the number of equal-width bins determined automatically using \package{numpy}'s {\tt bins='auto'}.
    {\it right:} a histogram created with \package{astropy}, with irregularly-spaced bins computed via the Bayesian Blocks algorithm.
    Compared to regularly-spaced bins, the irregular bin widths give a more accurate visual representation of features in the dataset at various scales.
    \label{fig:bayes-blocks-hist}
}
\end{figure}
